---
title: "Lab 4_Demo"
author: "Mateo Robbins"
date: "2024-04-22"
output: html_document
---

```{r packages, include = FALSE}
library(tidytext)
library(tidyverse)
library(tidymodels)
library(textrecipes)
library(discrim) # naive-bayes
library(naivebayes)
```

This data set includes more possible predictors than the text alone, but for this model we will only use the text variable
```{r data, message=FALSE, error=FALSE}

urlfile ="https://raw.githubusercontent.com/MaRo406/EDS-231-text-sentiment/main/data/climbing_reports_model_dat.csv"
incidents_df<-readr::read_csv(url(urlfile))
```

Now we'll split our data into training and test portions

```{r split-data, message=FALSE, error=FALSE}
set.seed(1234)

incidents2class <- incidents_df %>% 
  mutate(fatal = factor(ifelse(is.na(Deadly),
                               "non-fatal",
                               "fatal")))

# look at the distribution of fatal and non-fatal
table(incidents2class$fatal)

incidents_split <- initial_split(incidents2class, 
                                 strata = fatal) # samples equally from each group

incidents_train <- training(incidents_split)
incidents_test <- testing(incidents_split)

```

We use recipe() to specify the predictor and outcome variables and the data.

```{r recipe, error=FALSE}
incidents_rec <- recipe(fatal ~ Text, data = incidents_train)

recipe <- incidents_rec %>% 
  step_tokenize(Text) %>% 
  step_tokenfilter(Text, max_tokens = 1000) %>%  # only use 1000 words
  step_tfidf(Text)#coming from text recipes 

recipe
```

Next we add some familiar pre-processing steps on our Text variable: tokenize to word level, filter to the most common words, and calculate tf-idf.

```{r pre-process, message=FALSE, error=FALSE}

incidents_wf <- workflow() %>% 
  add_recipe(recipe)

```

Create  tidymodels workflow to combine the modeling components

```{r workflow, message=FALSE, error=FALSE}

nb_spec <- parsnip::naive_Bayes() %>% 
  set_mode("classification") %>% 
  set_engine("naivebayes") 

nb_spec 

```

```{r nb-spec, message=FALSE, error=FALSE}

nb_fit <- incidents_wf %>% 
  add_model(nb_spec) %>% 
  fit(data = incidents_train)

```

Now we are ready to add our model to the workflow and fit it to the training data

```{r fit-model, message=FALSE, error=FALSE}


```
Next up is model evaluation. We'll stretch our training data further and use resampling to evaluate our naive Bayes model. Here we create 10-fold cross-validation sets, and use them to estimate performance.

```{r cv_folds, message=FALSE, error=FALSE}
set.seed(999)
incidents_folds <- vfold_cv(incidents_train, 
                            v = 10) # default

```

```{r nb-workflow, message=FALSE, error=FALSE}

nb_wf <- workflow() %>% 
  add_recipe(recipe) %>% 
  add_model(nb_spec)
```

To estimate its performance, we fit the model to each of the resampled folds, and then evaluate on the heldout part of each resampled fold.

```{r fit-resamples, message=FALSE, error=FALSE}

nb_rs <- fit_resamples(
  nb_wf, 
  incidents_folds, 
  control = control_resamples(save_pred = T)
)

```

Extract the relevant information using collect_metrics() and collect_predictions() and examine the performance metrics.

```{r performance, message=FALSE, error=FALSE}

nb_rs_metrics <- collect_metrics(nb_rs)
nb_rs_predictions <- collect_predictions(nb_rs)

nb_rs_metrics
```

We'll use two performance metrics: accuracy and ROC AUC.
Accuracy is the proportion of the data that is predicted correctly. 
The ROC curve plots the true positive rate against the false positive rate; AUC closer to 1 indicates a better-performing model, while AUC closer to 0.5 indicates a model that does no better than random guessing.

```{r performance-plot, message=FALSE, error=FALSE}

nb_rs_predictions %>% 
  group_by(id) %>% 
  roc_curve(truth = fatal, 
            .pred_fatal) %>% 
  autoplot() +
  labs("Resampling",
       title = "ROC curve do climbing incident reports")

```

Another model method involves the confusion matrix. A confusion matrix tabulates a model's false positives and false negatives for each class.

```{r confusion-matrix, message=FALSE, error=FALSE}

conf_mat_resampled(nb_rs, tidy = F) %>% 
  autoplot(type = "heatmap")

```

```{r null-model, message=FALSE, error=FALSE}

```
